---
title: "final_755"
author: 'Madison Herrmann and Amanda Crosby'
date: "2024-08-01"
output: html_document
---


## Set up: Load libraries and packages

```{r}
#| include: false
library(tidyverse)
library(tidymodels)
library(ggplot2)

devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_eda.R?raw=true")
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/format_path.R?raw=true")
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/fun_plots.R?raw=true")
theme_set(theme_classic())
options(dplyr.print_max=Inf)

path_data <- "/Users/amandacrosby/755_final"
```

## Reading and cleaning names
```{r}
d <- read.csv(file.path(path_data, "heart_disease.csv")) |> janitor::clean_names() |> glimpse()
```

## Check for missing data and outliers
```{r}
d |> skim_all()
```

## Reclass outcome variable for classification
```{r}
d <- d |> 
  mutate(heart_disease = factor(heart_disease, levels = c(0, 1), labels = c("No", "Yes"))) |>   mutate(across(where(is.character), factor))
d |> skim_some()
```

## Distributions
Gender
```{r}
d |> tab(sex)
```

Age
```{r}
ggplot(d, aes(x = age)) +
  geom_histogram(binwidth = 5, fill = "blue", color = "black", alpha = 0.7) +
  labs(title = "Age Distribution",
       x = "Age",
       y = "Frequency") +
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5))
```


## Split data for EDA
```{r}
set.seed(20110522) 

splits <- d |> 
  initial_split(prop = .75, strata = "heart_disease")

splits |> 
  analysis() |> 
  glimpse() |> 
  write_csv(here::here(path_data, "hd_trn.csv"))

splits |> 
  assessment() |> 
  glimpse() |> 
  write_csv(here::here(path_data, "hd_val.csv"))
```

## EDA
Read in training data
```{r}
data_trn <- read_csv(here::here(path_data, "hd_trn.csv"),
                     col_type = cols()) |> 
  glimpse()
```

Reclass variables
```{r}
data_trn <- data_trn |> mutate(across(where(is.character), factor)) |> glimpse()
```

Outcome is balanced
```{r}
data_trn |> plot_bar("heart_disease")
```

Bivariate numeric with outcome
```{r}
data_trn |> 
  select(where(is.numeric)) |> 
  names() |> 
  map(\(name) plot_grouped_box_violin(df = data_trn, x = "heart_disease", y = name)) |> 
  cowplot::plot_grid(plotlist = _, ncol = 2)
```

Bivariate categorical with outcome
```{r}
data_trn |> 
  select(where(is.factor)) |>
  names() |>
  map(\(name) plot_bar(df = data_trn, x = name)) |>
  cowplot::plot_grid(plotlist = _, ncol = 2)
```

Correlation matrix 
```{r}
data_trn |> 
  mutate(heart_disease = as.numeric(heart_disease)) |> 
  select(where(is.numeric)) |> 
  cor(use = "pairwise.complete.obs") |> 
 corrplot::corrplot.mixed()
```

## Statistics

Reclass age as numeric

```{r}
d <- d |> 
  mutate(heart_disease = as.numeric(as.character(recode(heart_disease, "No" = 0, "Yes" = 1))))
d |> glimpse()
```

```{r}
m1 <- lm(heart_disease ~ age, d)
summary(m1)
```

There is a statistically significant relationship between age and heart disease. For every increase in age, the predicted probability of heart disease increases by 0.01, b = .01, F(1, 916) = 79.16, p = < .001.

```{r}
m2 <- lm(heart_disease ~ sex, d)
summary(m2)
```

There is a statistically significant relationship between sex and heart disease. Males were associated with a 0.37 increase in the predicted probability of heart disease, as opposed to females, b = .37, F(1, 916) = 94.25, p = < .001. 

```{r}
m3 <- lm(heart_disease ~ fasting_bs, d)
summary(m3)
```

There is a statistically significant relationship between fasting blood sugar and heart disease, b = .31, F(1, 916) = 70.48, p = < .001. A fasting blood sugar level over 120 mg/dL was associated with a 0.31 increase in the predicted probability of heart disease compared to having a normal blood sugar level.

```{r}
m4 <- lm(heart_disease ~ fasting_bs * sex * age, d)
summary(m4)
```

There was not a statistically significant interaction between fasting blood sugar and sex moderated by age, b =  0, F(7, 910) = 34.65, p = .78. 

## Graph
```{r}
m4_pred <- d |> 
  mutate(predicted = predict(m4, newdata = d))

ggplot(m4_pred, aes(x = age, y = predicted, color = factor(fasting_bs))) +
  geom_line() +
  facet_wrap(~ sex) +
  labs(title = "Interaction of Fasting Blood Sugar,\n Sex, and Age on Heart Disease",
       x = "Age",
       y = "Predicted Probability\n of Heart Disease",
       color = "Fasting BS") +
  scale_color_manual(values = c("red", "blue"), 
                     labels = c("≤ 120 mg/dL", "> 120 mg/dL")) +
  theme_bw() +
  theme(plot.title = element_text(size = 12, hjust = 0.5))
```


